---
title: "basic_tutorial"
author: "Bob Verity and Keith Fraser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
set.seed(1)
library(RMAPI)
```

The purpose of PlasmoMAPI is to use pairwise data on spatial and genetic
distances between populations to identify areas that are likely barriers or
corridors of gene flow. It does this by comparing the observed distribution of
relatedness to what we would expect under the null hypothesis of no systematic
spatial pattern in the data.

This vignette is designed to give a flavour of how PlasmoMAPI works and what
outputs it can produce. For a more detailed tutorial, including suggestions on
how to ensure your results are statistically valid, see the next vignette.

## Creating a project and loading data

PlasmoMAPI works using projects, which are essentially just lists containing all
data, input and outputs in one place. 

Next we need some data to work with. PlasmoMAPI requires two types of data: i)
spatial coordinates (longitude and latitude) of sampling locations, ii) pairwise
statistical values between all sampling locations - most likely estimates of
genetic relatedness. For this vignette we will use a dataset stored within the
package. These data were simulated under a simple model of genetics and
migration, and crucially with a *known barrier* and a *known corridor* of gene
flow that we will try to identify.

