---
title: "Basic tutorial"
author: "Bob Verity and Keith Fraser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basic_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE}
set.seed(1)
library(PlasmoMAPI)
```

*PlasmoMAPI* uses pairwise data on spatial and genetic distances to identify
areas that are likely barriers or corridors of gene flow. It does this by
comparing the observed distribution of genetic distances to what we would expect
under the null hypothesis of no systematic spatial pattern in the data.

This vignette is intended to give a flavour for how *PlasmoMAPI* works and what
outputs it can produce, without going into much detail. For a more in-depth
tutorial, including suggestions on how to ensure your results are statistically
valid, have a look at the next vignette.

## Creating a project and loading data

*PlasmoMAPI* works using projects, which are essentially just lists containing all
data, input and outputs in one place. We start by creating a new project:

```{r}
# create new project
p <- pm_project()
```

Now we need some data to work with. *PlasmoMAPI* requires two types of data: i)
spatial coordinates (longitude and latitude) of sampling locations, ii) pairwise
statistical values between all sampling locations. For this vignette we will use
a dataset stored within the package, which was simulated under a simple model of
*P. falciparum* genetics and epidemiology. Crucially this dataset was generated
with a **single known barrier** and a **single known corridor** of gene flow,
giving us something to aim for.

```{r}
# load tutorial data
tut_data <- pm_file("foo.rds")

# load sampling coordinates into project
p <- load_coords(p, tut_data$coords$long, tut_data$coords$lat)
```

Use the sampling coordinates to define a hex map. We then "assign" the map,
meaning we work out which pairwise values map to each hex, based on an
eccentricity value.

```{r, fig.width = 6, fig.height = 4.5}
# create and assign map
p <- create_map(p)
p <- assign_map(p, pb_markdown = TRUE)

# plot basic hex map
plot(p)
```

Load pairwise data, and produce a simple plot of spatial vs. statistical
distance. It is very hard to see any subtle trends from this plot, as everything
just looks like a cloud of data.

```{r, fig.width = 6, fig.height = 4.5}
# load pairwise genetic distances into project
p <- load_data(p, tut_data$genetic_distance)

# plot pairwise spatial vs. genetic distances
plot_dist(p)
```

Run the main *PlasmoMAPI* analysis.

```{r}
# run analysis
p <- pm_analysis(p, pb_markdown = TRUE)
```

Now produce the main output of the program - a map showing the z-score of every
hex compared to the null distribution. Areas that have significantly high/low
z-scores are outlined in black/white, and any known features are also overlaid
as dashed polygons. Note that the significance threshold takes into account
multiple testing, meaning under the null model we expect to see **no significant
hexes** 95% of the time. Hence, in this data there is actually a very strong
signal of a barrier to gene flow at the bottom, and a corridor of gene flow on
the right, which we know to be correct.

```{r, fig.width = 6, fig.height = 4.5}
# plot map
plot_map(p, barrier_list = tut_data$known_barriers)
```

We can also produce an interactive version of the same map:

```{r, fig.width = 6, fig.height = 4.5}
# plot interactive map
plot_leaflet(p)
```
