---
title: "Range-limited analysis"
author: "Bob Verity and Keith Fraser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Range-limited analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE}
set.seed(1)
library(PlasmoMAPI)
library(magrittr)
```

Sometimes barriers/corridors of gene flow can act at different spatial scales.
For example, a river might be a barrier to human movement when on foot but not
at longer distances because people tend to drive and so can quickly access the
nearest road bridge. We can explore this phenomenon in *PlasmoMAPI* by
restricting our analysis to edges of a certain length, and ignoring anything
outside the defined range. In some cases this can lead to sharper inference by
reducing the amount of random noise in the data.

## Simulating and loading data

We assume a uniform scatter of sampling locations, which we load into a new
project.

```{r}
# assume uniform scatter of sampling locations
node_df <- data.frame(long = runif(2e2, -10, 10),
                      lat = runif(2e2, -10, 10))

# create project and map
p <- pm_project() %>%
  load_coords(node_df$long, node_df$lat) %>%
  create_map() %>%
  assign_map(pb_markdown = TRUE)
```

As in the previous tutorial we will simulate some data with a known barrier to
gene flow using the `get_barrier_intersect()` function. This time we will assume
a more detailed barrier shape to test our ability to detect small and complex
features.

```{r, fig.width = 6, fig.height = 4.5}
# create barrier polygon
barrier_list <- list(data.frame(long = c(-5,0,6,7,4,-3,-2,3,4,3,-5,-5),
                                lat = c(5,7,6,0,-8,-4,-3,-4,-1,4,4,5)))

# plot map with barriers
plot_map(p, poly_list = barrier_list)
```

The argument `max_barrier_range` in `get_barrier_intersect()` allows us to
define a spatial distance beyond which the barrier penalty does not apply, for
example if `max_barrier_range = 100` and two sampling locations are either side
of the barrier and separated by 110km then the barrier effectively does not
exist between these locations. We will use this mtehod to simulate pairwise
genetic distances between sampling locations, this time on the [-Inf,Inf]
interval rather than the [0,1] interval just for a change.

```{r, fig.width = 6, fig.height = 4.5}
# calculate distances, taking into account barriers
sim_dist <- get_barrier_intersect(node_df$long, node_df$lat,
                                  barrier_list = barrier_list,
                                  barrier_penalty = -50,
                                  barrier_method = 2,
                                  max_barrier_range = 500)

# add some noise to the data
sim_dist_noise <- sim_dist + rnorm(length(sim_dist), sd = 2e2)
genetic_distance <- as.matrix(as.dist(sim_dist_noise))
```

Create a project, load the data, and look at a simple plot of spatial vs.
genetic distance.

```{r, fig.width = 6, fig.height = 4.5}
# load data and produce simple scatterplot
p <- load_data(p, genetic_distance)

plot_dist(p)
```

As before, it is quite difficult to see any systematic pattern in the data.


## Running a range-limited analysis

Now let's see what happens when we carry out the *PlasmoMAPI* analysis over the
entire range of data, and then a second time restricting to distances less than
500km:

```{r, fig.width = 6, fig.height = 4.5}
# run analysis over all distance and plot map
p <- pm_analysis(p, max_dist = Inf, pb_markdown = TRUE)
plot_map(p, poly_list = barrier_list)

# run analysis over short distance and plot map
p <- pm_analysis(p, max_dist = 500, pb_markdown = TRUE)
plot_map(p, poly_list = barrier_list)
```

Notice that we didn't find any areas of significance when using the whole
dataset because the random noise at distances above 500km was washing out any
signal, but when restricting to short distances the barrier becomes clear.
However, throwing data away like this comes at a cost - there are now
quite a number of hexes that are below our required coverage level, and so are
shown in grey. The more we subset the data, the fewer edges we we have at our
disposal and the fewer hexes will pass our coverage checks. So, performing a
focussed analysis like this is a trade-off between precision and statistical
power, and how far you can push it will depend on the number of sampling
locations at hand.
